---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
  word_document: default
  
---


### Loading and preprocessing the data
Change working directory and load the data.
```{r,echo=TRUE}
library(ggplot2)
setwd("~/Documents/Diverse/Coursera Reproducible Research /Peer Assesment 1/Peeras1/RepData_PeerAssessment1/")
df <- read.csv("activity.csv") #Read and convert to data.table
df$date.new <- strptime(paste(as.character(df$date),sprintf("%04.f",df$interval)),format = "%Y-%m-%d %H%M") #Create a new date column in format POSIX

```

### What is mean total number of steps taken per day?
Below is the code to produce a histogram on number of steps per day.
```{r,echo=TRUE}
df.perday <- aggregate(steps~date,df,sum)
qplot(steps,data=df.perday,geom="histogram",ylim=c(0,10),xlab="Number of steps per day",ylab="Count",binwidth=500)
```

Here the the code for producing hte mean and average.


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
